---
title: "RR-ProjectOne"
author: "Yon Hai"
date: "March 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data


```{r, results="hide"}
dir <- "C:/Users/Yonathan/Desktop/Coursera/data"
setwd(dir)

## packages 
 
library(ggplot2)
library(plyr)
library(lattice)
library(dplyr)

activity <- read.csv("activity.csv")

```

```{r}
names(activity)
str(activity)
View(activity)

summary(activity)
```

Transform factors into date format 

```{r}
activity$date <- as.POSIXct(activity$date, format="%Y-%m-%d")
```

### Total number of steps tacken each day 

We can create a collapsed dataset of aggrgated number of steps tacken per day 
```{r}
totalsteps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
colnames(totalsteps) <- c("Date", "Steps")
```

## Histogram of the total number of steps taken each day


```{r}
hist(totalsteps$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day", col = "blue")
```

## Mean and median number of steps taken each day

Mean 

```{r}

mean(totalsteps$Steps)
round (mean(totalsteps$Steps), digits = 0) 
```

Median 
```{r}
median(totalsteps$Steps)
```
 

##Average number of steps taken per interval 

```{r}
averageinterval <- with(activity, aggregate(steps, by = list(interval), FUN = mean, na.rm = TRUE))
colnames(averageinterval) <- c("interval", "meansteps")
```

##Time series plot of the average number of steps taken per interval 
```{r}
p <- ggplot(averageinterval, aes(x=interval, y=meansteps), xlab = "Interval", ylab="Average Daily Steps")
p + geom_line()+xlab("Interval")+ylab("Average Daily Steps")+ggtitle("Average Daily Steps per Interval")
```

## Imputation 

Total number of missing values 
```{r}
sum(is.na(activity$steps))
```
We can choose the values to impute missing data. This could be mean and median of the day or the interval. For the purpose of this project, I will take the mean of the interval to fill in the NA values  

```{r}

imputedsteps <- averageinterval$meansteps[match(activity$interval, averageinterval$interval)]
# create a new dataset, Always better to keep the orginal one
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputedsteps, no = activity$steps))

totalsteps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(totalsteps_imputed) <- c("date", "steps")
```
Now we can check the number of missing values 

sum(is.na(activity_imputed$steps))


## graph with the imputed ones 
```{r}
hist(totalsteps_imputed$steps, col = "red", xlab = "Steps", ylim = c(0,30), main = "Total number of steps taken each day", breaks = 5)
hist(totalsteps$Steps, col = "green", xlab = "Steps", ylim = c(0,30), main = "Total number of steps taken each day", breaks = 5, add=T)
legend("topright", c("Imputed Data", "Original Data"), fill=c("red", "green"))
```

## Differences in Activity between weekdays and weekends 

```{r}
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
    if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
    {y <- "Weekend"} else 
    {y <- "Weekday"}
    y
})
```

## Plot the difference 

```{r}
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
    geom_line() +
    labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
    facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```


